---
title: "Chapter 4 Exercises"
author: "Colin Gillespie"
output: rmarkdown::html_vignette
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Chapter 4 Exercises}
-->

### Exercise 1

First load the package and make a Spark connection

```{r eval=FALSE}
library("sparklyr")
sc = spark_connect(master = "local")
```

Then load the `dplyr` package
```{r eval=FALSE}
library("dplyr")
```

In this practical we're going to use the `nycflights13` data set. Run 
```{r eval=FALSE}
?nycflights13::flights
```
to get an overview of the data set.

To copy the data frame across to your Spark cluster, use the `copy_to` function
```{r eval=FALSE}
flights_tbl = copy_to(sc, nycflights13::flights, "flights")
```
List the available tables using `src_tbls(sc)`.

### Exercise 2

1. Using the `filter` dplyr function, select flights where the airtime was greater than 10 hours

    ```{r eval=FALSE}
    flights_tbl %>% filter(air_time > 10 * 60)
    ```


2. Lets investigate how delay varies with day of the month. 
The following code groups by 
day of the week, then works out the conditional arrival delay

    ```{r eval=FALSE}
    delay = flights_tbl %>% 
      group_by(day) %>%
      summarise(delay = mean(arr_delay))
    ```

    Now compare `delay` with
    ```{r eval=FALSE}
    delay_collect = delay %>% collect()
    ```
    What's the difference? We can the delays easily ggplot2 (or base graphics)
    ```{r eval=FALSE}
    library(ggplot2)
    ## Why does delay not work below?
    ggplot(delay_collect, aes(day, delay)) +
      geom_point() +
      geom_smooth()
    ```
3. Produce a similar plot for `dep_delay` variable.

4. **Tricky question**: Create a boxplot graphic where the x-axis is day of the week and the y-axis is the
delay.

    ```{r eval=FALSE, echo=FALSE}
    library(lubridate)
    delay$wday = wday(dmy(paste(delay$day, delay$month, delay$year)), label = T)
    ```

#### Links

  * [Spark](https://spark.apache.org/) homepage;
    * Latest [version](https://spark.apache.org/docs/latest/);
  * Stackoverflow: [sparkr](https://stackoverflow.com/questions/tagged/sparkr) & 
  [sparklyr]](https://stackoverflow.com/questions/tagged/sparklyr] tagged questions;
  * [SparkR](https://spark.apache.org/docs/latest/sparkr.html) & [sparklyr](http://spark.rstudio.com/) documentation
