---
title: "Chapter 3 Exercises"
author: "Jumping Rivers"
output: rmarkdown::html_vignette
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Chapter 3 Exercises}
-->


## Exercise 1

Load the `world_bank` dataset

```{r, results="hide", echo=-1, message=FALSE}
library("dplyr")
data(world_bank, package="jrBig")
```

Convert the data frame into a `dplyr` data frame
```{r}
wb = tbl_df(world_bank)
```

1. Print the `world_bank` and `wb` data frames to screen. What's different?
1. What does the `glimpse` function do? Hint: just try it on a data frame.
1. Using `wb`, complete the following tasks.
  * `filter`
    * Filter to retain data from the Country `AFG`;
    * Filter to retain data for years between 1980 and 1990.
  * `select`
    * Remove the `Year.Code` column;
    * Select columns `Year` and `gini`;
  * `mutate`
    * Create a new column called `Year2010` that is `yes` for rows where `Year > 2010`
  * `arrange`
    * Sort the data by `Year` in descending order and `gini`
  * `summarise`: 
    * Calculate the standard deviation of the `gdp_percap` column.
 * Bonus function. What does `slice` do? Try
    * `slice(wb, 1:3)`, `slice(wb, 5:10)`, `slice(wb, n())`
    * What does `n()` do?
1. Look at the documentation for the `sample_n` function. Can you sample
$100$ rows from the data set?

## Exercise 2

1. Calculate the mean `gini` and `gdp_percap` for each country; set `na.rm=TRUE` 
in the mean function. Hint: group by Country.
```{r echo=FALSE, results="hide"}
gb = group_by(wb, Year)
summarise(gb, mean(gini, na.rm=TRUE), mean(gdp_percap, na.rm=TRUE))
```
1. Calculate the median `gini` and `gdp_percap` for each country per year.
```{r results="hide", echo=FALSE}
gb = group_by(wb, Year, Country.Code)
summarise(gb, median(gini, na.rm=TRUE), median(gdp_percap, na.rm=TRUE))
```

Using the pipe operator, link the following operations together (for the `wb` data set)

1. Filter to retain data from the Country `AFG`;
    * then remove the `Year.Code` column;
    * then sort the data by `Year` in descending order and `gini`
1. Filter to retain data for years between 1980 and 1990;
    * then select columns `Year` and `gini`;
    * create a new column called `Year2010` that is `yes` for rows where `Year > 2010`
```{r echo=FALSE, results="hide"}
wb = tbl_df(world_bank)
wb %>% 
  filter(Country.Code == "AFG") %>%
  select(-1) %>%
  mutate(Year2010 = Year > 2010)
```

Compare 
    ```{r results="hide"}
    wb %>% group_by(Year, Country.Code) %>%
        summarise(gini=median(gini, na.rm=TRUE)) %>%
        summarise(max(gini, na.rm=TRUE))
    ```
and
    ```{r results="hide"}
      wb %>% group_by(Country.Code, Year) %>%
        summarise(gini=median(gini, na.rm=TRUE)) %>%
        summarise(max(gini, na.rm=TRUE))
    ```
  * Why are the answers different? What's happening?

## Exercise 3

1. Create an sql lite (or mysql, pgsql) database
    ```{r eval=FALSE}
    db = src_sqlite(path=tempfile(), create=TRUE)
    wb_sqlite = copy_to(db, world_bank, temporary = FALSE)
    wb_sqlite = tbl(db, "world_bank")
    src_desc(db) ## Gives you some details
    src_tbls(db) ## Lists the tables in the DB
    ```
1. Extract the top 50 rows from this table.
1. Redo exercise 3, but using the data base query. 
    * Examine the underlying SQL code;
    * Use `collect` to get the database.


---
**Tip**: Check out [dplyr](https://cran.r-project.org/web/packages/dplyr/)'s CRAN 
page.