---
title: "Timing Reading and Writing Files"
author: "Colin Gillespie"
output: rmarkdown::html_vignette
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Timing Reading and Writing Files}
-->

## Example data set

In the following examples, we'll use a data set that has one million rows and
ten columns. The associated data frame is $800$MB. To make things interesting,
the data frame will include a mixture of numbers and missing values.

```{r eval=FALSE}
x = runif(1e7)
x[sample(1e7, 1e6)] = NA # 10% NAs
dd =  as.data.frame(replicate(10, x))
@


```{r include=FALSE, echo=FALSE, cache=TRUE}
x = runif(1e3)
x[sample(1e3, 1e2)] = NA # 10% NAs
dd =  as.data.frame(replicate(10, x))
```

### Timings

Base R: `read.csv`/`write.csv`

```{r results="hide", cache=TRUE}
system.time(write.csv(dd, "dd.csv"))
#   user  system elapsed 
#101.012   1.096 104.153 
system.time(read.csv("dd.csv"))
#   user  system elapsed 
#151.568   0.728 152.134 
```

Base R: `readRDS`/`saveRDS`
  
```{r results="hide", cache=TRUE}
system.time(saveRDS(dd, file="dd.RData"))
#   user  system elapsed 
# 101.77    0.28  102.69 
system.time(readRDS(file="dd.RData"))
#   user  system elapsed 
#  5.052   0.104   5.148 
````

The [data.table](https://github.com/Rdatatable/data.table/wiki) package

```{r results="hide", cache=TRUE}
library("data.table")
system.time(fwrite(dd, "dd.csv"))
#   user  system elapsed 
# 28.776   1.600   4.128 
system.time(fread("dd.csv"))
#   user  system elapsed 
#  21.93    0.18   22.07 
```
The [readr](https://github.com/hadley/readr) package}
```{r results="hide", cache=TRUE}
library("readr")
system.time(write_csv(dd, "dd.csv"))
#   user  system elapsed 
# 189.44    1.54  190.77 
system.time(read_csv("dd.csv"))
#   user  system elapsed 
# 18.476   0.252  18.588 
```

The [feather](https://github.com/wesm/feather) package
```{r results="hide", cache=TRUE, eval=FALSE}
library("feather")
system.time(write_feather(dd, "dd.feather"))
#   user  system elapsed 
#  0.620   0.540   1.319 
system.time(read_feather("dd.feather"))
#   user  system elapsed 
#  0.356   0.068   0.422 
```
Personally, I think that `feather` looks very promising, but is very new.

### Overview

```{r results="hide", echo=FALSE, fig.cap="Overall comparison of read/write times for an $800$MB file.", fig.width=5}
writes = c(104, 102, 4.1, 190, 1.319)
reads = c(152, 5.1, 22.07, 18.58, 0.422)
n = c("read.csv", "Rdata", "data.table", "readr", "feather")

par(mar=c(3,3,2,1), mgp=c(2,0.4,0), tck=-.01,
    cex.axis=0.8, las=1, xaxs='i',yaxs='i')
barplot(writes, names.arg=n, col="steelblue", 
        border=NA, space=0.05, ylim=c(0, 200), 
        ylab="Time(s)")
abline(h=seq(0, 200, 20), col="white")
title("Write times", adj=1, 
      cex.main=1, font.main=2, col.main="black")

 par(mar=c(3,3,2,1), mgp=c(2,0.4,0), tck=-.01,
     cex.axis=0.8, las=1, xaxs='i',yaxs='i')
barplot(reads, names.arg=n, col="steelblue", 
        border=NA, space=0.05, ylim=c(0, 200), 
        ylab="Time(s)")
abline(h=seq(0, 150, 20), col="white")
title("Read times", adj=1, 
      cex.main=1, font.main=2, col.main="black")
```