---
title: "Chapter 3 Solutions"
author: "Jumping Rivers"
output: rmarkdown::html_vignette
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Chapter 3 solutions}
-->

## Exercise 1

Load the `world_bank` dataset

```{r, results="hide", echo=-1, message=FALSE}
library("dplyr")
data(world_bank, package = "jrBig")
```

Convert the data frame into a `dplyr` data frame
```{r}
wb = tbl_df(world_bank)
```

1. Print the `world_bank` and `wb` data frames to screen. What's different?
    ```{r results="hide"}
    ## wb only prints out 10 rows & and gives types
    ```
1. What does the `glimpse` function do? Hint: just try it on a data frame.
    ```{r}
    ## Flips the data frame on it's side
    ```
1. Using `wb`, complete the following tasks.
  * `filter`
    * Filter to retain data from the Country `AFG`;
    * Filter to retain data for years between 1980 and 1990.
    ```{r results="hide"}
    filter(wb, Country.Code == "AFG")
    filter(wb, Year > 1980 & Year < 1990)
    ```
  * `select`
    * Remove the `Year.Code` column;
    * Select columns `Year` and `gini`;
    ```{r results="hide"}
    select(wb, -1)
    select(wb, Year, gini)
    ```
  * `mutate`
    * Create a new column called `Year2010` that is `yes` for rows where `Year > 2010`
    ```{r results="hide"}
    mutate(wb, Year2010 = ifelse(Year > 2010, "yes", "no"))
    ```
  * `arrange`
    * Sort the data by `Year` in descending order and `gini`
    ```{r results="hide"}
    arrange(wb, desc(Year), gini)
    ```
  * `summarise`: 
    * Calculate the standard deviation of the `gdp_percap` column.
    ```{r results="hide"}
    summarise(wb, mean(gdp_percap, na.rm = TRUE))
    ```
 * Bonus function. What does `slice` do? Try
    * `slice(wb, 1:3)`, `slice(wb, 5:10)`, `slice(wb, n())`
    * What does `n()` do? __Counts__
1. Look at the documentation for the `sample_n` function. Can you sample
$100$ rows from the data set?

## Exercise 2

1. Calculate the mean `gini` and `gdp_percap` for each country; set `na.rm=TRUE` 
in the mean function. Hint: group by Country.
```{r echo=FALSE, results="hide"}
gb = group_by(wb, Year)
summarise(gb, mean(gini, na.rm = TRUE), mean(gdp_percap, na.rm = TRUE))
```
1. Calculate the median `gini` and `gdp_percap` for each country per year.
```{r results="hide", echo=FALSE}
gb = group_by(wb, Year, Country.Code)
summarise(gb, median(gini, na.rm = TRUE), median(gdp_percap, na.rm = TRUE))
```
1. Compare 
    ```{r results="hide"}
    wb %>%
      group_by(Year, Country.Code) %>%
      summarise(gini = median(gini, na.rm = TRUE)) %>%
      summarise(max(gini, na.rm = TRUE))
    ```
and
    ```{r results="hide"}
      wb %>%
      group_by(Country.Code, Year) %>%
      summarise(gini = median(gini, na.rm = TRUE)) %>%
      summarise(max(gini, na.rm = TRUE))
    ```
  * Why are the answers different? What's happening? 
    * __Solution__ With each application of `summarise` a variable is peeled off the
    group_by statement.

## Exercise 3

Using the pipe operator, link the following operations together (for the `wb` data set)

1. Filter to retain data from the Country `AFG`;
    * then remove the `Year.Code` column;
    * then sort the data by `Year` in descending order and `gini`
    ````{r eval=FALSE}
    wb %>%
      select(Country == "AFG") %>%
      arrange(desc(YEAR), gini)
    ```
1. Filter to retain data for years between 1980 and 1990;
    * then select columns `Year` and `gini`;
    * create a new column called `Year2010` that is `yes` for rows where `Year > 2010`

    ```{r echo=FALSE, results="hide"}
    wb %>% 
      filter(Year > 1980 & Year < 1990) %>%
      select(-1) %>%
      mutate(Year2010 = ifelse(Year > 2010, "yes", "no"))
    ```


## Exercise 4

1. Create an sql lite (or mysql, pgsql) database
    ```{r eval=FALSE}
    db = src_sqlite(path = tempfile(), create = TRUE)
    wb_sqlite = copy_to(db, world_bank, temporary = FALSE)
    wb_sqlite = tbl(db, "world_bank")
    src_desc(db) ## Gives you some details
    src_tbls(db) ## Lists the tables in the DB
    ```
1. Extract the top 50 rows from this table.
    ```{r eval=FALSE}
    head(wb_sqlite, 50)
    ```
1. Redo exercise 3, but using the data base query. 
    * Examine the underlying SQL code;
    * Use `collect` to get the database.


---
**Tip**: Check out [dplyr](https://cran.r-project.org/web/packages/dplyr/)'s CRAN  page.
