---
title: "Rcpp Solutions"
author: "Colin Gillespie"
output: rmarkdown::html_vignette
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Rcpp Solutions}
-->


## Overview

The exercises below have been written assuming that you have never programmed in 
C/C++ before. If you have previous experience, feel free to alter the questions to suit
your background.

## Exercise 1

1. Load the `Rcpp` package 
    ```{r, message=FALSE}
    library("Rcpp")
    ```
and run the following code
    ```{r eval=FALSE}
    cppFunction('
      double add_c(double x, double y){
        double value = x + y;
        return value;
      }
    ')
```
1. Call the function `add_c` and make sure it works. What happens if you pass it a vector
or a character to the function?
1. Hard: Change the function to calculate $x^y$. Hint: Google taking powers in C++.

```{r engine="Rcpp", eval=FALSE}
 cppFunction('
      double pow_c(double x, double y){
        double value = pow(x, y);
        return value;
      }
    ')
```

## Exercise 2

1. Create a directory called `src` - this is a standard directory name for computer code.
It's short for `source`. This is where you should keep your C++ code.
1. Create a file called `add.cpp` and move the `add_c` function (with necessary headers)
to the file.
    * Compile the file and call the `add_c` function.
1. Similar to above, but add a function called `subtract_c`.

## Exercise 3

1. Write an R function that calculates the sample standard deviation - you are **not** 
allowed to use `sum`, `mean`, `sd` or `var`, i.e. Use a for loop. The formula for the standard deviation 
is

$$
\sigma = \sum_{i=0}^n (x_i - \mu)
$$
where $\mu$ is the mean. __Hint__ you will need a function to calculate
the mean, i.e. `mean_r`

```{r results="hide"}
mean_r = function(x) {
  n = length(x)
  m = 0
  for(i in seq_along(x)) 
    m = m + x[i]/n
  m
}

sd_r = function(x) {
  n = length(x)
  var = 0; m = mean_r(x)
  for(i in seq_along(x)) {
    var = var + (x[i] - m)^2
  }
  sqrt(var/(n-1))
}
x = rnorm(10)
mean_r(x); mean(x)
sd_r(x); sd(x)
```
1. Write a C function that calculates the standard deviation. 
```{r engine="Rcpp", eval=FALSE}
## Should be in a file
double mean_c(NumericVector x){
  int i;
  int n = x.size();
  double mean = 0;

  for(i=0; i<n; i++) {
    mean = mean + x[i]/n;
  }
  return mean;
}
double sd_c(NumericVector x){
  int i;
  int n = x.size();
  double mean = mean_c(x);
  double var=0;
  for(i=0; i<n; i++) {
    var = var + pow(x[i] - mean, 2);
  }
  return pow(var/(n-1), 0.5);
}
```

1. Using `microbenchmark`, compare your C version to your R function.
    - Use the base R function `sd` as a baseline.


-----
Hint: There's a chapter on RCpp [`Efficient R programming`](https://csgillespie.github.io/efficientR/efficient-performance.html#rcpp)




