---
title: "Chapter 4 Exercises"
author: "Colin Gillespie"
output: rmarkdown::html_vignette
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Chapter 4 Exercises}
-->

### Exercise 1

Load the package and make a connection

```{r eval=FALSE}
library(sparklyr)
sc = spark_connect(master = "local")
```
In this exercise, we're going to use the `nycflights13` data set. Look at
the associated help page `?nycflights13::flights` to get an overview of the data set.
Load the `dplyr` package
```{r eval=FALSE}
library(dplyr)
```
and copy the dataframe across
```{r eval=FALSE}
flights_tbl = copy_to(sc, nycflights13::flights, "flights")
```
List the available tables using `src_tbls(sc)`. Look

### Exercise 2

1. Using the `filter` dplyr function, select flights where the airtime was greater than 10 hours

```{r eval=FALSE}
flights_tbl %>% filter(air_time > 10*60)
```


2. Lets investigate how delay varies with day of the month. The following code groups by 
day of the week, then works out the conditional arrivial delay

```{r eval=FALSE}
delay = flights_tbl %>% 
  group_by(day) %>%
  summarise(delay = mean(arr_delay))
```

Now compare
```{r eval=FALSE}
delay
delay_collect = delay %>% collect()
```
What's the difference? We can the delays easily ggplot2 (or base graphics)
```{r eval=FALSE}
ggplot(delay, aes(day, delay)) +
  geom_point() +
  geom_smooth()
```
3. Produce a similar plot for `dep_delay`


4. **Tricky question**: Create a boxplot graphic where the x-axis is day of the week and the y-axis is the
delay.


```{r eval=FALSE}
library(lubridate)
delay$wday = wday(dmy(paste(delay$day, delay$month, delay$year)), label = T)
```

#### Links

  * [Spark](https://spark.apache.org/) homepage;
    * Latest [version](https://spark.apache.org/docs/latest/);
  * Stackoverflow: [sparkr](https://stackoverflow.com/questions/tagged/sparkr) & 
  [sparklyr]](https://stackoverflow.com/questions/tagged/sparklyr] tagged questions;
  * [SparkR](https://spark.apache.org/docs/latest/sparkr.html) & [sparklyr](http://spark.rstudio.com/) documentation
  
### Tasks

  * There is a nice section on creating dataframes in the [SparkR](https://spark.apache.org/docs/latest/sparkr.html) documentation.
  * Also have a look at the machine learning section.
    