---
title: "Rcpp Exercises"
author: "Colin Gillespie"
output: rmarkdown::html_vignette
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Rcpp}
-->
  
## Overview

The exercises below have been written assuming that you have never programmed in 
C/C++ before. If you have previous experience, feel free to alter the questions to suit
your background.

## Exercise 1

1. Load the `Rcpp` package 
    ```{r, message=FALSE}
    library("Rcpp")
    ```
and run the following code
    ```{r eval=FALSE}
    cppFunction('
      double add_c(double x, double y){
        double value = x + y;
        return value;
      }
    ')
```
1. Call the function `add_c` and make sure it works. What happens if you pass it a vector
or a character to the function?
1. Hard: Change the function to calculate $x^y$. Hint: Google taking powers in C++.

## Exercise 2

1. Create a directory called `src` - this is a standard directory name for computer code.
It's short for `source`. This is where you should keep your C++ code.
1. Create a file called `add.cpp` and move the `add_c` function (with necessary headers)
to the file.
    * Compile the file and call the `add_c` function.
1. Similar to above, but add a function called `subtract_c`.

## Exercise 3

1. Write an R function that calculates the standard deviation - you are **not** 
allowed to use `sum`, `mean`, `sd` or `var`.
1. Write a C function that calculates the standard deviation. 
1. Using `microbenchmark`, compare your C version to your R function.
    - Use the base R function `sd` as a baseline.


-----
Hint: There's a chapter on RCpp [`Efficient R programming`](https://csgillespie.github.io/efficientR/efficient-performance.html#rcpp)









