---
title: "Chapter 5: The Rcpp Package"
author: Colin Gillespie
date: "`r Sys.Date()`"
output: ioslides_presentation
css: left.css
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Chapter 5 Slides}
-->

# 5.1 Introduction
## 5.1 Introduction

  * Sometimes R is just slow. 
  * You've tried every trick you know, and your code is still crawling along
  * Rewrite code in C/C++
    * Using `.Call` is very painful

## Rcpp

  * There's a better way! Rcpp!
     * In this chapter, C and C++ code are largely [interchangeable](http://programmers.stackexchange.com/q/16390/14846)
  * Goal: provide a flavour of what's possible.
    * No previous C++ experience

## Exercise: Are you ready?

```{r eval=FALSE}
install.packages("Rcpp")
```

```{r results="hide"}
jrBig::test_rcpp()
```
See Pre-requisites in the notes if you have problems

# 5.2 A simple C++ function
## 5.2 A simple C++ function

```{r}
## R version
add_r = function(x, y) {
  return(x + y)
} 
```

```{r engine='Rcpp', eval=FALSE}
/* Return type double
 * Two arguments, also doubles
 */
double add_c(double x, double y) {
    double value = x + y;
    return value;
}
```

> Exercise: What differences do you see?

## The `main` function

  * Typically we need a function `main`; not here. 
  * Rcpp makes it easy to call C++ functions from R
  * Users doen't have to worry about
    * platform;
    * compilers;
    * R/C interface.

## The `cppFunction` command

```{r message=FALSE}
library("Rcpp")
cppFunction('
  double add_c(double x, double y){
    double value = x + y;
    return value;
  }
')
```

Rcpp then performs magic!

## The magic

```{r}
add_c
```

We can call the `add_c`} function in the usual way

```{r}
add_c(1, 2)
```

# Chapter 3: Exercise 1
## Chapter 3: Exercise 1

```{r eval=FALSE}
vignette("Rcpp", package="jrBig")
```

# 5.2.2 C/C++ data types
## 5.2.2 C/C++ data types

 Type | Description
------|-----------
`char` | A single character.
`int` | An integer.
`float` | A single precision floating point number.
`double` | A double-precision floating point number.
`void` | A valueless quantity.

There are also `pointer`'s


# 5.3 The sourceCpp function
## 5.3 The sourceCpp function

  * The `cppFunction` is great for small examples
  * But you should put C++ code in a separate file (`.cpp`)
    * `sourceCpp("path/to/file.cpp")`
  * The `.cpp` needs a few headers
  
## `.Cpp` components

Access Rcpp functions (similar to a `library` call)
```{r engine='Rcpp', eval=FALSE}
#include <Rcpp.h>
```

## Namespace


  * To access the Rcpp functions we would have to type `Rcpp::function_1`
  * To avoid typing `Rcpp::`
    ```{r engine="Rcpp", eval=FALSE}
    using namespace Rcpp;
    ```
  * Above each function we want to export/use in R
    ```{r engine="Rcpp", eval=FALSE}
    // [[Rcpp::export]]
    ```

## Complete file

```{r engine="Rcpp"}
# include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
double add_c(double x, double y) {
  double value = x + y;
  return value;
}
```

## Benefits of a file

1. Syntax highlighting
1. It's easy to make errors switching between R and C++ in the same file

To save space, we we'll omit the headers for the remainder of the chapter.

# Exercise 2
## Exercise 2
```{r eval=FALSE}
vignette("Rcpp", package="jrBig")
```

# 5.4 Vectors and loops
## `mean_r`

```{r}
mean_r = function(x) {
  n = length(x)
  m = 0
  for(i in seq_along(x))
    m = m + x[i]/n
  m
}
```

## `mean_c`

```{r engine="Rcpp",eval=FALSE}
double mean_c(NumericVector x){
  int i;
  int n = x.size();
  double mean = 0;

  for(i=0; i<n; i++) {
    mean = mean + x[i]/n;
  }
  return mean;
}
sourceCpp("../src/mean_c.cpp")
```
```{r echo=FALSE} 
cppFunction('double mean_c(NumericVector x){
  int i;
  int n = x.size();
  double mean = 0;

  for(i=0; i<n; i++) {
    mean = mean + x[i]/n;
  }
  return mean;
}')
```

## Speed comparison

```{r}
library("microbenchmark")
```

We generate some normal random numbers for the comparison

```{r}
x = rnorm(1e4)
```

Then call the `microbenchmark` function.

```{r cache=TRUE}
z = microbenchmark(
  mean(x),
  mean_r(x),
  mean_c(x)
)
```

## Results

```{r echo=FALSE}
par(mar=c(3,3,2,1), mgp=c(2,0.4,0), tck=-.01, cex.axis=0.9, las=1)
boxplot(z, ylab="Time", col="steelblue")
grid()
```

# Exercise 3
## Exercise 3
```{r eval=FALSE}
vignette("Rcpp", package="jrBig")
```


# 5.5 C++ with sugar on top
## 5.5 The sugared version

  * Rcpp sugar brings a higher-level of abstraction to C++ code written using the
Rcpp API. 

```{r engine="Rcpp", eval=FALSE}
NumericVector res_c(NumericVector x, NumericVector y) {
  int i;
  int n = x.size();
  NumericVector residuals(n);
  for(i=0; i<n; i++) {
    residuals[i] = pow(x[i] - y[i], 2);
  }
  return residuals;
}
```

## With Sugar

```{r engine="Rcpp", eval=FALSE}
NumericVector res_sugar(NumericVector x, NumericVector y) {
  return pow(x - y, 2);
}
```

  * The sugar versions aren't usually faster than the C++ version, 
     - but there's usually very little difference between the two. 
  * However with the sugared variety, the code is shorter and is constantly being improved.
