---
title: "Graphics Solutions"
author: "Colin Gillespie"
output: rmarkdown::html_vignette
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Graphics solutions}
-->

## Exercise 1

1. Check that you have the `bigvis` package installed.
    ```{r, message=FALSE}
    ## Should run error free
    library("bigvis")
    data("movies", package="bigvis")
    ```
1. Use `condense` and `bin` to plot the movie lengths. Try varying the `width` argument
in the `bin` function.
1. Add a smoothed line using the `smooth` function.
1. Load the `ggplot2` package
    ```{r message=FALSE}
    library("ggplot2")
    ```
Use `autoplot` on the binned data. Experiment with `peel`.

```{r eval=FALSE, echo=FALSE}
x = condense(bin(movies$length, width = 1))
plot(x)
sm = smooth(x, h = 5 / 100)
lines(sm[,1], sm[,2], col=2)
autoplot(x)
last_plot() %+% peel(x)
```

## Exercise 2

1. Check that you have the `tabplot` package installed.
    ```{r, message=FALSE}
    ## Should run with no errors
    library("tabplot")
    ```

1. Reproduce the plot in the notes
    ```{r eval=FALSE}
    tableplot(movies[,3:5])
    ```
Use `head(movies[,3:5])` to examine the first few rows of the data. What do the plots
mean?

1. Add the argument `sortCol=3` to the above `tableplot` command. What does this argument
do?

1. Run the following command 

    ```{r eval=FALSE}
    tableplot(movies[,3:10])
    ```
What do the graphs mean?

---
Did you know that we also run on-site courses?
